Comparing bigstatsr and bigsnpr with PLINK and EIGENSOFT

### 2018-01-14 compare-GWAS-PRS-desktop.Rmd

```{r}
suppressMessages({
  # devtools::install_github("privefl/bigsnpr")
  library(bigsnpr)   
  # install.packages("glue")
  library(glue)      
  # install.packages("ggplot2")
  library(ggplot2)
})
# FastPCA: https://data.broadinstitute.org/alkesgroup/EIGENSOFT/EIG-6.1.4.tar.gz
smartpca <- "../EIG-6.1.4/bin/smartpca"
plink <- download_plink()

@@

ind.keep2 <- match(scan("plink.prune.in", what = ""), celiac$map$marker.ID)
mean(ind.keep %in% ind.keep2)

@@

# PRS
(thrs <- c(0, exp(seq(log(0.2), log(550), length.out = 50))))
lpS <- -predict(gwas.train)
system.time(
  prs <- snp_PRS(G, betas.keep = gwas.train$estim[ind.keep],
                 ind.test = ind.test,
                 ind.keep = ind.keep,
                 lpS.keep = lpS[ind.keep], 
                 thr.list = thrs)
)

# AUC
aucs <- apply(prs, 2, AUC, target = y01[ind.test])
```

@@

```{r}
# Get a file with only training set
write.table2(cbind(celiac$fam[ind.train, 1:2], 
                   obj.svd$u[ind.train, ]), "covar_train.txt")
# GWAS on training set
system.time(
  system(glue("{plink} --bfile {prefix_bed}",
              " --keep covar_train.txt",
              " --logistic hide-covar beta",
              " --covar covar_train.txt",
              " --threads {NCORES}"))
)

@@

clump <- fread2("plink.clumped")
ind.keep2 <- match(clump$SNP, celiac$map$marker.ID)
mean(ind.keep %in% ind.keep2)

@@

scores <- matrix(NA_real_, length(ind.test), length(thrs))
for (i in seq_along(thrs)) {
  scores[, i] <- fread2(glue("plink.S{i}.profile"))$SCORESUM
}

#AUC
aucs2 <- apply(scores, 2, AUC, target = y01[ind.test])

# Number of predictor for each threshold
nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
bigstatsr:::MY_THEME(qplot(nb.pred, aucs, color = "A", log = "x", 
                           cex = I(3), pch = I(1),
                           xlab = "#predictors", ylab = "AUC"), coeff = 2) +
  geom_line(show.legend = FALSE) + 
  geom_point(aes(nb.pred, aucs2, color = "B"), cex = 3, pch = 3) + 
  scale_colour_manual(
    name = NULL, 
    values = c("black", "red"),
    labels = c("PRS with bigstastr\nand bigsnpr", "PRS with plink")) + 
  guides(color = guide_legend(override.aes = list(shape = c(1, 3)))) + 
  theme(legend.position = c(0.7, 0.2)) 
```

>>>>

```{r}
suppressMessages({
  # devtools::install_github("privefl/bigsnpr")
  library(bigsnpr)   
  # install.packages("glue")
  library(glue)      
  # install.packages("ggplot2")
  library(tidyverse)
})
# FastPCA: https://data.broadinstitute.org/alkesgroup/EIGENSOFT/EIG-6.1.4.tar.gz
smartpca <- "../EIG-6.1.4/bin/smartpca"
plink <- download_plink()

@@

ind.keep2 <- match(scan("plink.prune.in", what = ""), celiac$map$marker.ID)
length(intersect(ind.keep, ind.keep2)) / length(union(ind.keep, ind.keep2))

@@

# PRS
(thrs <- c(0, exp(seq(log(0.1), log(100), length.out = 100))))
lpS <- -predict(gwas.train)
system.time(
  prs <- snp_PRS(G, betas.keep = gwas.train$estim[ind.keep],
                 ind.test = ind.test,
                 ind.keep = ind.keep,
                 lpS.keep = lpS[ind.keep], 
                 thr.list = thrs)
)

# AUC
aucs <- apply(prs, 2, AUC, target = y01[ind.test])
# Number of predictor for each threshold
nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
```

@@

```{r, fig.width = 10}
# Get a file with only training set
write.table2(cbind(celiac$fam[ind.train, 1:2], 
                   obj.svd$u[ind.train, ]), "covar_train.txt")
# GWAS on training set
system.time(
  system(glue("{plink} --bfile {prefix_bed}",
              " --keep covar_train.txt",
              " --logistic hide-covar beta",
              " --covar covar_train.txt",
              " --threads {NCORES}"))
)

@@

clump <- fread2("plink.clumped")
ind.keep2 <- match(clump$SNP, celiac$map$marker.ID)
length(intersect(ind.keep, ind.keep2)) / length(union(ind.keep, ind.keep2))

@@

scores2 <- matrix(NA_real_, length(ind.test), length(thrs))
nb.pred2 <- numeric(length(thrs))
for (i in seq_along(thrs)) {
  tmp <- fread2(glue("plink.S{i}.profile"))
  scores2[, i] <- tmp$SCORESUM
  nb.pred2[i]  <- tmp$CNT[[1]] / 2
}

#AUC
aucs2 <- apply(scores2, 2, AUC, target = y01[ind.test])

bind_rows(
  data_frame(nb = nb.pred,  auc = aucs,  Software = "bigsnpr"),
  data_frame(nb = nb.pred2, auc = aucs2, Software = "PLINK")
) %>%
  ggplot() %>%
  bigstatsr:::MY_THEME() +
  geom_point(aes(nb, auc, color = Software, shape = Software), size = 5) +
  geom_line(aes(nb, auc, color = Software, linetype = Software), show.legend = FALSE) +
  scale_x_log10() + 
  scale_colour_manual(
    values = c("black", "red"),
    labels = c("PRS with bigstastr\nand bigsnpr", "PRS with plink")) + 
  scale_shape_manual(
    values = c(1, 3),
    labels = c("PRS with bigstastr\nand bigsnpr", "PRS with plink")) +
  labs(x = "#predictors", y = "AUC") +
  theme(legend.position = c(0.6, 0.2)) 
```
